---
title: "Binomial Distributions"
subtitle: "A Discrete Probability Distribution"
author: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: ["default","default-fonts","https://svgsstats.github.io/Styles/my_theme.css"]
    #css: ["default"] #,"tamu"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "custom.js"
    chakra: "remark-latest.min.js"
---
```{r echo=FALSE, error=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
```
# Warmup

The mean cost of textbooks for a class at a university is $176 with a standard deviation of $62.  Assuming that textbooks prices are normally distributed, find the following: 

1. What portion of classes have textbooks that cost more than $200? 

2. A class whose textbook is in the top 5% of all textbook costs has textbooks that cost more than what value? 

3. The middle 80% of textbook costs are between what two values? 
---
# Warmup

The mean cost of textbooks for a class at a university is $176 with a standard deviation of $62.  Assuming that textbooks prices are normally distributed, find the following: 

(1)  What portion of classes have textbooks that cost more than $200? 

`r 1-pnorm(200,176,62)`


(2)  A class whose textbook is in the top 5% of all textbook costs has textbooks that cost more than what value? 

`r qnorm(0.95,176,62)`


(3)  The middle 80% of textbook costs are between what two values? 

```{r echo=FALSE}
qnorm(0.10,176,62)
qnorm(0.9,176,62)
```
---
# Continuous vs Discrete Random Variables

<p class="inverse border white tab">
Discrete random variables are random variables that take on a countable list of values
</p>

For example: 
- Number of heads a coin flip gets in 50 flips. 

- Number of individuals with a particular disease in a sample of size 1000. 
---
# Continuous vs Discrete Random Variables
<p class="inverse border white tab">
Continuous random variables are random variables which take on infinitely many values</p> 

For example: 
- How much time it takes for popcorn to finish popping

- The weight of a randomly selected labradoodle. 
<img src="https://vetstreet.brightspotcdn.com/dims4/default/7d8a842/2147483647/thumbnail/645x380/quality/90/?url=https%3A%2F%2Fvetstreet-brightspot.s3.amazonaws.com%2F1c%2F854d80a8b511e0a0d50050568d634f%2Ffile%2Flabradoodle-2-645mk070111.jpg" width="50%"></img>
---
# Normal distributions are one distribution that describes some continuous random variables. 

<br/><br/><br/>

# Binomial distributions are one distribution that describes some discrete random variables. 
---
# What is a Binomial Setting? 

A **binomial setting** arises when we perform several independent trials of the same chance process and record the number of times that a particular outcome occurs.

The four conditions are: 
- **Binary** - The possible outcomes of each trial can be classified as "success" or "failure".
- **Independent** - Trials must be independent.  Knowing the outcome of one trial must not have any effect on the result of any other trial. 
- **Number** - The number of trials *n* must be fixed in advance. 
- **Success** - The probability of success *p* must be the same for every trial. 

--

# BINS!
---
# Is X a Binomial random variable? 

- Observe number of heads when flipping a coin 50 times.  X = number of heads

---
# Is X a Binomial random variable? 

- Observe sex of next 50 children born at a local hospital.   X = number of girls

---
# Is X a Binomial random variable? 

- Ten-question quiz has five True/False questions, and five multiple-choice questions each with four possible choices.  Student randomly picks an answer for every questions.   X = number of answers that are correct

---
# Is X a Binomial random variable? 

- Four students are randomly picked without replacement from a student body containing 35 juniors and 35 seniors.  X = number of seniors among the four selected students. 

---
# Is X a Binomial random variable? 

- A die rolled until the number 5 is chosen.  X = number of rolls it takes before a 5 is rolled. 

---
# Motivating Example

Each child of a particular pair of parents has probability 0.25 of having type O blood. Mendel said that children receive genes from each of their parents independently. If these parents have 5 children, the count X of children with type O blood is a binomial random variable with n = 5 trials and probability p = 0.25 of a success on each trial. In this setting, a child with type O blood is a “success” (S) and a child with another blood type is a “failure” (F). 

What’s P(X = 2)?

--

P(SSFFF) = (0.25)(0.25)(0.75)(0.75)(0.75) = 0.02637

--

However, there are a number of different arrangements in which 2 out of the 5 children have type O blood: 

SSFFF SFSFF SFFSF SFFFS FSSFF
FSFSF FSFFS FFSSF FFSFS FFFSS

--

There are 10 different arrangements, each with probability of 0.02637 of occurring.  So, $P(X = 2) = 10(0.25)^2(0.75)^3 = 0.2637$
---
# Technology

Just like with normal distribution, we can use R to speed up the calculation. 

```{r}
dbinom(x=2,size=5,prob=0.25)
```

---
# R Commands in Binomial Distributions

```{r eval=FALSE}
dbinom(x,size,prob)
```
This command will give the probability of getting exactly *x* successes in *size* trials with probability of success of *prob*. 

```{r eval=FALSE}
pbinom(q,size,prob)
```
This command will give the probability of getting less than or equal to *q* successes in *size* trials with probability of success of *prob*. 

```{r eval=FALSE}
qbinom(p,size,prob)
```
This command will give you the smallest *x* value such that P(X >= x) is greater than *p*. Put another way, the number of successes below which at least p% of options lie. 

---
# Some Examples

Each child of a particular pair of parents has probability 0.25 of having blood type O. Suppose the parents have 5 children

(a) Find the probability that exactly 3 of the children have type O blood.

--
```{r}
dbinom(x=3,size=5,prob=0.25)
```


---
# Some Examples

Each child of a particular pair of parents has probability 0.25 of having blood type O. Suppose the parents have 5 children

(b) Should the parents be surprised if more than 3 of their children have type O blood?

--

```{r}
1 - pbinom(q=3,size=5,prob=0.25)
```


---
# SOL Guessing?

"I guessed on the entire SOL and still passed," says hypothetical student. 

We wish to determine the probability that a student could actually do this. A previous version of the Algebra 1 SOL had 50 multiple choice questions and a student needed to score a 54% to pass it (I hope this angers you).  There were 4 options for each question.


(a)  If a student flat out guessed on all 50 questions, what are the chances that they could pass the SOL? 

--
```{r}
0.54*50

1 - pbinom(26,50,0.25)
```

---
# SOL Guessing?

"I guessed on the entire SOL and still passed," says hypothetical student. 

We wish to determine the probability that a student could actually do this. A previous version of the Algebra 1 SOL had 50 multiple choice questions and a student needed to score a 54% to pass it (I hope this angers you).  There were 4 options for each question.


(b)  What if the student was able to eliminate just one answer on each of the 50 questions, what are the chances that they could pass the SOL?

--
```{r}
1 - pbinom(26,50,1/3)
```


---
# SOL Guessing?

"I guessed on the entire SOL and still passed," says hypothetical student. 

We wish to determine the probability that a student could actually do this. A previous version of the Algebra 1 SOL had 50 multiple choice questions and a student needed to score a 54% to pass it (I hope this angers you).  There were 4 options for each question.


(c)  What if they could eliminate two answers? 

--
```{r}
1 - pbinom(26,50,0.5)
```


---
# SOL Guessing?

"I guessed on the entire SOL and still passed," says hypothetical student. 

We wish to determine the probability that a student could actually do this. A previous version of the Algebra 1 SOL had 50 multiple choice questions and a student needed to score a 54% to pass it (I hope this angers you).  There were 4 options for each question.


(d)  What is the range of correct answers that the middle 95% of test-takers would get on the Algebra 1 SOL (assuming that they guessed)

--
```{r}
qbinom(0.025,50,0.25)
qbinom(0.975,50,0.25)
```


---
# Another Example

A blind-folded marksman finds that on average he hits the target 80% of the time.  If he fires 4 shots, find the probability that he has

(a) more than 2 hits?  

--
```{r}
1 - pbinom(2,4,0.80)
```


(b) at least 3 misses? 

--
```{r}
1 - pbinom(2,4,0.20)
```


---
# Characteristics of the Binomial Distribution

Let's consider the SOCS of the Binomial Distributions. 

First, let's simulate 10,000 sequences of 100 coin flips
```{r}
simulate <- rbinom(n=10000, size=100, prob=0.5) 
```

---
# Center
The mean of the distribution represents the long-term average number of heads we would see if we repeated the 100 coin flips many many times.     


```{r eval=FALSE}
mean(simulate)
```

--

```{r echo=FALSE}
mean(simulate)
```

--

$$ \mu_x = n \cdot p $$
---
# Spread - Standard Deviation

The standard deviation of the distribution represents the typical distance between observations of the random variable and the mean of the random variable. 


```{r eval=FALSE}
sd(simulate)
```

--

```{r echo=FALSE}
sd(simulate)
```

--

$$ \sigma_x = \sqrt{n \cdot p \cdot (1 - p)} $$
---
# Shape

```{r echo=FALSE}
library(tidyverse)
library(ggthemes)
```

```{r fig.width=5, fig.height=4}
qplot(simulate,binwidth=1) + theme_light()
```

--

It is true that binomial distributions SOMETIMES look approximately normal.  This is not true always.  There are certain conditions that must be met for this to be true. 

---
# Shape
```{r fig.width=5,fig.height=4}
simulate <- rbinom(n=10000,size = 20, prob = 0.05)
qplot(simulate,binwidth=1) + theme_light()
```

---
# Shape

A *binomial distribution* is approximately normal if **there are at least 10 expected successes and failures**. 

We will check: 
    $$np >= 10$$
    $$n(1-p) >= 10$$

---
# Bottled Water vs Tap Water

Suppose that 21 students were involved in an experiment in which they did a blind taste test with 3 different types of water.  One of the water types was tap, and the other two were different bottled waters.   Let X = number of students who could correctly identify the tap water.   Let us assume that the students just flat out guessed on which one was the tap water. 

Find the mean of x. 

--

$$ \mu_x = np $$
$$ \mu_x = 21(\frac{1}{3}) = 7 $$

--

We would expect, on average, 7 students to guess correctly. 

---
# Bottled Water vs Tap Water

Suppose that 21 students were involved in an experiment in which they did a blind taste test with 3 different types of water.  One of the water types was tap, and the other two were different bottled waters.   Let X = number of students who could correctly identify the tap water.   Let us assume that the students just flat out guessed on which one was the tap water. 

Find the standard deviation of x. 

--

$$ \sigma_x = \sqrt{np(1-p)} $$

$$ \sigma_x = \sqrt{21(\frac{1}{3})(1-\frac{1}{3})} = 2.16 $$



--

If the activity were repeated many times with groups of 21 students who were just guessing, the number of correct identifications would differ from 7 by an average of 2.16.

---
# Attitudes Toward Shopping

A survey asked a nationwide random sample of 2500 adults if they agreed or disagreed that “I like buying new clothes, but shopping is often frustrating and time-consuming.” Suppose that exactly 60% of all adult US residents would say “Agree” if asked the same question. Let X = the number in the sample who agree. 

Estimate the probability that 1520 or more of the sample agree.

1. Verify that X is approximately a binomial random variable.

2. Check the conditions for using a Normal approximation.

3. Calculate P(X ≥ 1520) using a Normal approximation.

---
# Note about Independence

We can assume independence even in cases that aren't truly independent as long as the sample isn't more than 10% of the population size. 

