---
title: "Binomial Distributions"
subtitle: "A Discrete Probability Distribution"
author: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: ["default","default-fonts","https://svgsstats.github.io/Styles/my_theme.css"]
    #css: ["default"] #,"tamu"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "custom.js"
    chakra: "remark-latest.min.js"
---
```{r echo=FALSE, error=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
```
# Warmup

The mean cost of textbooks for a class at a university is $176 with a standard deviation of $62.  Assuming that textbooks prices are normally distributed, find the following: 

1. What portion of classes have textbooks that cost more than $200? 

2. A class whose textbook is in the top 5% of all textbook costs has textbooks that cost more than what value? 

3. The middle 80% of textbook costs are between what two values? 
---
# Warmup

The mean cost of textbooks for a class at a university is $176 with a standard deviation of $62.  Assuming that textbooks prices are normally distributed, find the following: 

(1)  What portion of classes have textbooks that cost more than $200? 

`r 1-pnorm(200,176,62)`


(2)  A class whose textbook is in the top 5% of all textbook costs has textbooks that cost more than what value? 

`r qnorm(0.95,176,62)`


(3)  The middle 80% of textbook costs are between what two values? 

```{r echo=FALSE}
qnorm(0.10,176,62)
qnorm(0.9,176,62)
```
---
# Continuous vs Discrete Random Variables

<p class="inverse border white tab">
Discrete random variables are random variables that take on a countable list of values
</p>

For example: 
- Number of heads a coin flip gets in 50 flips. 

- Number of individuals with a particular disease in a sample of size 1000. 
---
# Continuous vs Discrete Random Variables
<p class="inverse border white tab">
Continuous random variables are random variables which take on infinitely many values</p> 

For example: 
- How much time it takes for popcorn to finish popping

- The weight of a randomly selected labradoodle. 
<img src="https://vetstreet.brightspotcdn.com/dims4/default/7d8a842/2147483647/thumbnail/645x380/quality/90/?url=https%3A%2F%2Fvetstreet-brightspot.s3.amazonaws.com%2F1c%2F854d80a8b511e0a0d50050568d634f%2Ffile%2Flabradoodle-2-645mk070111.jpg" width="50%"></img>
---
# Normal distributions are one distribution that describes some continuous random variables. 

<br/><br/><br/>

# Binomial distributions are one distribution that describes some discrete random variables. 
---
# What is a Binomial Setting? 

A **binomial setting** arises when we perform several independent trials of the same chance process and record the number of times that a particular outcome occurs.

The four conditions are: 
- **Binary** - The possible outcomes of each trial can be classified as "success" or "failure".
- **Independent** - Trials must be independent.  Knowing the outcome of one trial must not have any effect on the result of any other trial. 
- **Number** - The number of trials *n* must be fixed in advance. 
- **Success** - The probability of success *p* must be the same for every trial. 

--

# BINS!
---
# Is X a Binomial random variable? 

- Observe number of heads when flipping a coin 50 times.  X = number of heads

---
# Is X a Binomial random variable? 

- Observe sex of next 50 children born at a local hospital.   X = number of girls

---
# Is X a Binomial random variable? 

- Ten-question quiz has five True/False questions, and five multiple-choice questions each with four possible choices.  Student randomly picks an answer for every questions.   X = number of answers that are correct

---
# Is X a Binomial random variable? 

- Four students are randomly picked without replacement from a student body containing 35 juniors and 35 seniors.  X = number of seniors among the four selected students. 

---
# Is X a Binomial random variable? 

- A die rolled until the number 5 is chosen.  X = number of rolls it takes before a 5 is rolled. 

---
# Motivating Example

Each child of a particular pair of parents has probability 0.25 of having type O blood. Mendel said that children receive genes from each of their parents independently. If these parents have 5 children, the count X of children with type O blood is a binomial random variable with n = 5 trials and probability p = 0.25 of a success on each trial. In this setting, a child with type O blood is a “success” (S) and a child with another blood type is a “failure” (F). 

What’s P(X = 2)?

--

P(SSFFF) = (0.25)(0.25)(0.75)(0.75)(0.75) = 0.02637

--

However, there are a number of different arrangements in which 2 out of the 5 children have type O blood: 

SSFFF SFSFF SFFSF SFFFS FSSFF
FSFSF FSFFS FFSSF FFSFS FFFSS

--

There are 10 different arrangements, each with probability of 0.02637 of occurring.  So, $P(X = 2) = 10(0.25)^2(0.75)^3 = 0.2637$
---
# Technology

Just like with normal distribution, we can use R to speed up the calculation. 

```{r}
dbinom(x=2,size=5,prob=0.25)
```

---
# R Commands in Binomial Distributions

```{r eval=FALSE}
dbinom(x,size,prob)
```
This command will give the probability of getting exactly *x* successes in *size* trials with probability of success of *prob*. 

```{r eval=FALSE}
pbinom(q,size,prob)
```
This command will give the probability of getting less than or equal to *q* successes in *size* trials with probability of success of *prob*. 

```{r eval=FALSE}
qbinom(p,size,prob)
```
This command will give you the smallest *x* value such that P(X >= x) is greater than *p*. Put another way, the number of successes below which at least p% of options lie. 

---
# Some Examples

Each child of a particular pair of parents has probability 0.25 of having blood type O. Suppose the parents have 5 children

(a) Find the probability that exactly 3 of the children have type O blood.

--
```{r}
dbinom(x=3,size=5,prob=0.25)
```


---
# Some Examples

Each child of a particular pair of parents has probability 0.25 of having blood type O. Suppose the parents have 5 children

(b) Should the parents be surprised if more than 3 of their children have type O blood?

--

```{r}
1 - pbinom(q=3,size=5,prob=0.25)
```


---
# SOL Guessing?

"I guessed on the entire SOL and still passed," says hypothetical student. 

We wish to determine the probability that a student could actually do this. A previous version of the Algebra 1 SOL had 50 multiple choice questions and a student needed to score a 54% to pass it (I hope this angers you).  There were 4 options for each question.


(a)  If a student flat out guessed on all 50 questions, what are the chances that they could pass the SOL? 

--
```{r}
0.54*50

1 - pbinom(26,50,0.25)
```

---
# SOL Guessing?

"I guessed on the entire SOL and still passed," says hypothetical student. 

We wish to determine the probability that a student could actually do this. A previous version of the Algebra 1 SOL had 50 multiple choice questions and a student needed to score a 54% to pass it (I hope this angers you).  There were 4 options for each question.


(b)  What if the student was able to eliminate just one answer on each of the 50 questions, what are the chances that they could pass the SOL?

--
```{r}
1 - pbinom(26,50,1/3)
```


---
# SOL Guessing?

"I guessed on the entire SOL and still passed," says hypothetical student. 

We wish to determine the probability that a student could actually do this. A previous version of the Algebra 1 SOL had 50 multiple choice questions and a student needed to score a 54% to pass it (I hope this angers you).  There were 4 options for each question.


(c)  What if they could eliminate two answers? 

--
```{r}
1 - pbinom(26,50,0.5)
```


---
# SOL Guessing?

"I guessed on the entire SOL and still passed," says hypothetical student. 

We wish to determine the probability that a student could actually do this. A previous version of the Algebra 1 SOL had 50 multiple choice questions and a student needed to score a 54% to pass it (I hope this angers you).  There were 4 options for each question.


(d)  What is the range of correct answers that the middle 95% of test-takers would get on the Algebra 1 SOL (assuming that they guessed)

--
```{r}
qbinom(0.025,50,0.25)
qbinom(0.975,50,0.25)
```


---
# Another Example

A blind-folded marksman finds that on average he hits the target 80% of the time.  If he fires 4 shots, find the probability that he has

(a) more than 2 hits?  

--
```{r}
1 - pbinom(2,4,0.80)
```


(b) at least 3 misses? 

--
```{r}
1 - pbinom(2,4,0.20)
```



